Characterization and Photometric Performance of the Hyper Suprime-Cam Software Pipeline.

### 2017-04-28 draft/ms_v5.tex
PSF color

%% ------------------------------------------------------------------------------------ %%
%% Fig: Magnitude v.s S/N for stars
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_sn}
    \end{center}
    \caption{
        Relation between input magnitudes (\textbf{left}: $i$-band; \textbf{right}:
        $g$-band) of synthetic stars and the $\log (S/N)$ measured by \hscpipe{} PSF
        photometry.
        Hexagonal binned density plots and contours show the distributions for
        stars in both the good- and bad-seeing \tracts{}. 
        Highly blended stars are highlighted using scatter plots.
        The gray dashed line marks $S/N = 5$ which is the detection threshold used 
        by \hscpipe{}.  
        Our 5$\sigma$ point source detection limit is ${\sim}26.5$ mag in $i$-band,
        however we do not reach these magnitudes with our current synthetic stellar 
        input catalog.
        }
    \label{fig:star_sn}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Accuracy of Forced PSF Magnitudes
\begin{figure*}
    \begin{center}
        \includegraphics[width=16cm]{fig/synpipe_psf_mag}
    \end{center}
    \caption{
        Accuracy of the \texttt{hscpipe} PSF photometry for synthetic stars measured
        by the difference between input and output \forced{} PSF magnitudes.
        Plots [a, b, c, d, e] show the results for [$g$, $r$, $i$, $z$, $y$]-band, 
        respectively.
        The left panel in each plot shows the relation between input magnitude and
        magnitude difference.
        Hexagonal binned density plots and contours are for the synthetic galaxies 
        from good and bad seeing \tracts{}.
        Highly blended objects are highlighted using scattered plots.
        The long-dashed lines mark zero magnitude difference, while the pairs of
        dashed lines outline the running-median of PSF magnitude errors
        (including the uncertainties in aperture correction).
        The right panel in each plot shows the distributions of the magnitude 
        differences for objects in good (filled) and bad (solid line) seeing \tracts{}.
        The dashed lines identify the distribution of highly blended objects.
        }
    \label{fig:psf_mag}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Difference between the Unforced and Forced Magnitudes
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_diff}
    \end{center}
    \caption{
        The magnitude differences between the \unforced{} and \forced{}
        PSF photometry for synthetic stars in $g$- (\textbf{left} and $y$-band
        (\textbf{right}).
        The lines and contours legend is identical to the plots in Fig \ref{fig:psf_mag}.
        }
    \label{fig:psf_diff}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Accuracies of the PSF color
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_color}
    \end{center}
    \caption{
        Accuracy of the color measurements for synthetic stars via the differences
        between input and \forced{} PSF colors.
        The \textbf{upper panels} and \textbf{lower panels} are for $g-i$ and $i-y$
        colors separately.
        The \textbf{left} column shows the relation between input magnitude and
        the color difference, and the \textbf{right} column uses the input colors as
        x--axis instead.
        The lines and contours legend is identical to the plot in Fig \ref{fig:psf_mag}.
        }
    \label{fig:psf_color}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    The precision of PSF magnitude for the \texttt{badSeeing} \tract{} shows similar 
    performance at $i_{\mathrm{Input}}<23.5$ mag. 
    At fainter magnitudes, the precision slightly degrades. 
    The \texttt{badSeeing} \tract{} has ${\sim}5$\% precision at
    $i_{\mathrm{Input}}{\sim}24.0$ mag and ${\sim}11$\% precision at
    $i_{\mathrm{Input}}{\sim}25.2$ mag.
    \citet{HSCDR1} evaluates the precision of PSF magnitude via external comparisons
    with the PS1, PV2, and SDSS data at $i<21$ mag, and finds it at 1-2\% level,
    which is consistent with our results. 
    However, at fainter magnitudes, external comparisons become difficult due to the 
    lack of imaging matched to HSC depths.
    The results from \synpipe{} hence provide useful evaluations of the precision of 
    the HSC photometry down to our detection limit, though, strictly speaking, the 
    precision reported here should be considered as upper limits because 
    \synpipe{} currently does not consider the systematic uncertainties in PSF 
    modeling.

@@

	Finally, we also perform similar tests for the \unforced{} PSF photometry in all 
	five bands, and find similar results. 
	To measure the \forced{} PSF photometry, \hscpipe{} fixes the centroid of the PSF 
	model across all five bands. 
	A difference between \forced{} and \unforced{} PSF magnitudes would therefore be 
	an indication of photometric uncertainties arising from  inaccurate astrometric 
	calibrations and PSF modeling across different bands. 
	Figure \ref{fig:psf_diff} shows the difference between \forced{} and \unforced{} 
	PSF photometry for the $g$ and the $y$ band. 
	The overall differences are small ($<0.01$ mag) .

@@

    In Fig \ref{fig:psf_color}, we evaluate the accuracy of the \forced{} $(g-i)$
    and $(i-y)$ PSF colors by comparing the error in color measurements with the input magnitudes colors. Figure  \ref{fig:psf_color} shows that \hscpipe{} provides accurate and unbiased PSF color
    estimates for synthetic stars with realistic color distributions down to very 
    faint magnitudes. The average uncertainty for $(g-i)$ is  ${\sim}0.025$ mag for $i_{\mathrm{Input}}{\sim}19.0$. This uncertainty increases to ${\sim}0.11$ mag at
    $i_{\mathrm{Input}}{\sim}24.0$ mag.  The $(i-y)$ color displays similar uncertainties. Figure \ref{fig:psf_color} also shows that the accuracy of the PSF colors do not depend on the input colors of the synthetic stars or the seeing conditions. For instance, the accuracy of $(g-i)$ for the bad-seeing \tract{} is only
    slightly worse in the very faint end (${\sim}0.14$ mag at
    $i_{\mathrm{Input}}{\sim}24.0$ mag) compared to the 
    good-seeing \tract{}.

Figure \ref{fig:psf_cdist} displays the uncertainties in PSF colors  by comparing the color distributions of the input synthetic star sample to the recovered color distributions. Using four different colors and two color--color planes, we show that the 
    \forced{} PSF colors accurately recover the distributions of all four colors 
    and the color-color distributions.

    As for highly blended stars, Fig \ref{fig:psf_color} also shows that highly blended stars have larger uncertainties in their colors compared to isolated stars. At fixed input magnitude, the  uncertainty in $(g-i)$ is ${\sim}0.1$ mag larger for highly blened stars, Also, for blended stars, $(g-i)$ shows a bias towards bluer colors, while $(i-y)$ shows a bias towards redder colors. However, such biases appear to be less severe in terms of colors than in terms of magnitudes.

@@

%% ------------------------------------------------------------------------------------ %%
%% Fig: Magnitude v.s S/N for galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_sn}
    \end{center}
    \caption{
        Relation between input magnitudes (\textbf{left}: $i$-band; \textbf{right}:
        $g$-band) of synthetic galaxies $log(S/N)$ as measured by \hscpipe{} \cmodel{} photometry. Lines and contours are similar to Fig \ref{fig:star_sn}
        }
    \label{fig:cmodel_sn}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Accuracy of Forced cModel Magnitudes for galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=16cm]{fig/synpipe_galaxy_mag}
    \end{center}
    \caption{
        Accuracies of the \texttt{hscpipe} \cmodel{} photometry for synthetic
        galaxies measured by the difference between input and output \forced{}
        \cmodel{} magnitudes.
        Plots [a, b, c, d, e] show the results for [$g$, $r$, $i$, $z$, $y$]-bands, 
        respectively.
        The lines and contours legend is identical to Fig \ref{fig:psf_mag}.
        }
    \label{fig:cmodel_mag}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Difference between the Unforced and Forced Magnitudes
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_cmodel_diff}
    \end{center}
    \caption{
        The magnitude differences between the \unforced{} and \forced{}
        \cmodel{} photometry for synthetic galaxies in $g$- (\textbf{left} and
        $y$-band (\textbf{right}).
        The lines and contours legend is identical to the plots in Fig \ref{fig:psf_diff}.
        }
    \label{fig:cmodel_diff}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    For galaxies, \hscpipe{} uses a \cmodel{} photometry algorithm that is an improved 
    version of the SDSS \cmodel{} photometry.
    For more details, please see Bosch\etal (in prep.).
    In short, \cmodel{} tries fits object using both a de~Vaucouleurs and an exponential profile.  The \cmodel{} algorithm then fits a linear combination of these two best fit models. Despite the limitations of the \cmodel{} method (e.g., sensitivity to the background
    subtraction and to deblending failures), it can deliver robust PSF-corrected fluxes and
    colors for galaxies.

    We randomly inject ${\sim}40,000$ synthetic galaxies into each
    \tract{}; and, using an approach similar to that for synthetic stars, we select 
    galaxy samples for photometric comparisons. Synthetic galaxies that are located within 2 pixels 
    of the centroids of real objects are removed from the sample  (these represent ${\sim}3$--$7$\% of the input sample). After imposing quality cuts on the  recovered synthetic galaxy sample (see Appendix \ref{app:qc} for details), we have ${\sim}30,000$
    galaxies in each \tract{}.

>>>>

%% ------------------------------------------------------------------------------------ %%
%% Fig: Magnitude v.s S/N for stars
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_sn}
    \end{center}
    \caption{
        Relation between input magnitudes (\textbf{left}: $i$-band; \textbf{right}:
        $g$-band) of synthetic stars and the $\log (\mathrm{S}/\mathrm{N})$ measured 
        by \hscpipe{} PSF photometry.
        Hexagonal binned density plots and contours show the distributions for
        stars in both the good- and bad-seeing \tracts{}. 
        Highly blended stars are highlighted using scatter plots.
        The gray dashed line marks $S/N = 5$ which is the detection threshold used 
        by \hscpipe{}.  
        Our 5$\sigma$ point source detection limit is ${\sim}26.5$ mag in $i$-band,
        however we do not reach these magnitudes with our current synthetic stellar 
        input catalog.
        }
    \label{fig:star_sn}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Accuracy of Forced PSF Magnitudes
\begin{figure*}
    \begin{center}
        \includegraphics[width=16cm]{fig/synpipe_psf_mag}
    \end{center}
    \caption{
        Accuracy of the \texttt{hscpipe} PSF photometry for synthetic stars measured
        by the difference between input and output \forced{} PSF magnitudes.
        Plots [a, b, c, d, e] show the results for [$g$, $r$, $i$, $z$, $y$]-band, 
        respectively.
        The left panel in each plot shows the relation between input magnitude and
        magnitude difference.
        Hexagonal binned density plots and contours are for the synthetic galaxies 
        from good and bad seeing \tracts{}.
        Highly blended objects are highlighted using scattered points.
        The long-dashed lines mark zero magnitude difference, while the pairs of
        dashed lines outline the running-median of PSF magnitude errors
        (including the uncertainties in aperture correction).
        The right panel in each plot shows the distributions of the magnitude 
        differences for objects in good (filled) and bad (solid line) seeing \tracts{}.
        The dashed lines identify the distribution of highly blended objects.
        }
    \label{fig:psf_mag}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Difference between the Unforced and Forced Magnitudes
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_psf_diff}
    \end{center}
    \caption{
        The magnitude differences between the \unforced{} and \forced{}
        PSF photometry for synthetic stars in $g$ (\textbf{left}) and $y$ band
        (\textbf{right}).
        The lines and contours legend is identical to the plots in Fig \ref{fig:psf_mag}.
        }
    \label{fig:psf_diff}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Accuracies of the PSF color
\begin{figure*}
    \begin{center}
        \includegraphics[width=15cm]{fig/synpipe_psf_color}
    \end{center}
    \caption{
        Accuracy of the color measurements for synthetic stars via the differences
        between input and \forced{} PSF colors.
        The \textbf{upper panels} and \textbf{lower panels} are for $g-i$ and $i-y$
        colors separately.
        The \textbf{left} column shows the relation between input magnitude and
        the color difference, and the \textbf{right} column uses the input colors as
        x--axis instead.
        The lines and contours legend is identical to the plot in Fig \ref{fig:psf_mag}.
        }
    \label{fig:psf_color}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    The precision of PSF magnitude for the \texttt{badSeeing} \tract{} shows similar 
    performance at $i_{\mathrm{Input}}<23.5$ mag. 
    At fainter magnitudes, the precision slightly degrades. 
    The \texttt{badSeeing} \tract{} has ${\sim}5$\% precision at
    $i_{\mathrm{Input}}{\sim}24.0$ mag and ${\sim}11$\% precision at
    $i_{\mathrm{Input}}{\sim}25.2$ mag.
    \citet{HSCDR1} evaluates the precision of PSF magnitude via external comparisons
    with the PS1, PV2, and SDSS data at $i<21$ mag, and finds it at 1-2\% level,
    which is consistent with our results. 
    However, at fainter magnitudes, external comparisons become difficult due to the 
    lack of imaging matched to HSC depths.
    The results from \synpipe{} hence provide useful evaluations of the precision of 
    the HSC photometry down to our detection limit\footnote{Strictly speaking, the 
    precision reported here should be considered as upper limits because 
    \synpipe{} currently does not consider the systematic uncertainties in PSF 
    modeling. But this effect should be subdominant.}

@@

	Finally, we also perform similar tests for the \unforced{} PSF photometry in all 
	five bands, and find similar results. 
	To measure the \forced{} PSF photometry, \hscpipe{} fixes the centroid of the PSF 
	model across all five bands. 
	A difference between \forced{} and \unforced{} PSF magnitudes would therefore be 
	an indication of photometric uncertainties arising from inaccurate astrometric 
	calibrations and PSF modeling across different bands\footnote{On average, the 
	centroids are more accurate in \forced{} photometry as they are defined using 
	the band with higher \s2n{}}. 
	Figure \ref{fig:psf_diff} shows the difference between \forced{} and \unforced{} 
	PSF photometry for the $g$ and the $y$ band. 
	The overall differences are small ($<0.01$ mag). 
	At very faint end ($g_{\mathrm{Input}}>24.0$ mag and $g_{\mathrm{Input}}>23.5$

@@

    In Fig \ref{fig:psf_color}, we evaluate the precision and accuracy of the 
    \forced{} $(g-i)$ and $(i-y)$ PSF colors by comparing the differences between input 
    and output colors ($\Delta \mathrm{Color}$) with both the input magnitudes and 
    colors. 
    Figure \ref{fig:psf_color} shows that \hscpipe{} provides precise and accurate PSF 
    colors for synthetic stars with realistic color distributions down to very 
    faint magnitudes. 
    The average precision for $(g-i)$ is ${\sim}0.025$ mag at 
    $i_{\mathrm{Input}}{\sim}19.0$. 
    The scatter increases to ${\sim}0.11$ mag at $i_{\mathrm{Input}}{\sim}24.0$ mag.  
    The $(i-y)$ color displays similar precision. 
    The \forced{} PSF color measurements are not biased through the entire ranges of 
    input magnitudes and colors.

    Figure \ref{fig:psf_color} shows that the precision of PSF colors do not depend 
    the seeing conditions. 
    For instance, the precision of $(g-i)$ for the \texttt{badSeeing} \tract{} is 
    only slightly worse at the very faint end (${\sim}0.14$ mag at
    $i_{\mathrm{Input}}{\sim}24.0$ mag) compared to the \texttt{goodSeeing} \tract{}.

    Figure \ref{fig:psf_cdist} displays the precision of PSF colors by comparing the 
    color distributions of the input sample to the recovered color distributions. 
    Using four different colors and two color--color planes, we show that the 
    \forced{} PSF colors accurately recover the distributions of all four colors 
    and the color-color distributions.

    As for highly blended stars, Fig \ref{fig:psf_color} also shows that highly blended 
    stars have worse precision and accuracy in their colors compared to isolated stars. 
    At fixed input magnitude, the precision in $(g-i)$ is ${\sim}0.1$ mag worse for 
    highly blended stars.  
    Also, for blended stars, $(g-i)$ shows a bias towards bluer colors, while $(i-y)$ 
    shows a bias towards redder colors. 
    However, such biases appear to be less severe in terms of colors than in terms of
    PSF magnitudes.

@@

%% ------------------------------------------------------------------------------------ %%
%% Fig: Magnitude v.s S/N for galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_galaxy_sn}
    \end{center}
    \caption{
        Relation between input magnitudes (\textbf{left}: $i$-band; \textbf{right}:
        $g$-band) of synthetic galaxies $\log (\mathrm{S}/\mathrm{N})$ as measured 
        by \hscpipe{} \cmodel{} photometry. 
        Lines and contours are similar to Fig \ref{fig:star_sn}
        }
    \label{fig:cmodel_sn}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Accuracy of Forced cModel Magnitudes for galaxies
\begin{figure*}
    \begin{center}
        \includegraphics[width=16cm]{fig/synpipe_galaxy_mag}
    \end{center}
    \caption{
        Accuracies of the \hscpipe{} \cmodel{} photometry for synthetic
        galaxies measured by the difference between input and output \forced{}
        \cmodel{} magnitudes.
        Plots [a, b, c, d, e] show the results for [$g$, $r$, $i$, $z$, $y$]-bands, 
        respectively.
        The lines and contours legend is identical to Fig \ref{fig:psf_mag}.
        }
    \label{fig:cmodel_mag}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
%% Fig: Difference between the Unforced and Forced Magnitudes
\begin{figure*}
    \begin{center}
        \includegraphics[width=\textwidth]{fig/synpipe_cmodel_diff}
    \end{center}
    \caption{
        The magnitude differences between the \unforced{} and \forced{}
        \cmodel{} photometry for synthetic galaxies in $g$ (\textbf{left}) and
        $y$ band (\textbf{right}).
        The lines and contours legend is identical to the plots in Fig 
        \ref{fig:psf_diff}.
        }
    \label{fig:cmodel_diff}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    For galaxies, \hscpipe{} uses a \cmodel{} photometry algorithm that is an improved 
    version of the SDSS \cmodel{} photometry.
    For details about the \cmodel{} algorithm, please see description in Section 
    \ref{ssec:hscpipe} and Bosch\etal (in prep.).
    Despite the limitations of the \cmodel{} method (e.g., sensitivity to the 
    background subtraction and to deblending failures), it can deliver robust 
    PSF-corrected fluxes and colors for galaxies.

    Typical \tract{} in the Wide layer contains ${\sim}400000$ extended objects with 
    $i_{\mathrm{CModel}} < 25.5$ mag. 
        We randomly inject ${\sim}40,000$ synthetic galaxies into each \tract{}.

    so we inject additional ${\sim}10$\% extended 
    objects to the images. 
    Using an approach similar to that for synthetic stars, we select 
    galaxy samples for photometric comparisons. Synthetic galaxies that are located within 2 pixels 
    of the centroids of real objects are removed from the sample  (these represent ${\sim}3$--$7$\% of the input sample). After imposing quality cuts on the  recovered synthetic galaxy sample (see Appendix \ref{app:qc} for details), we have ${\sim}30,000$
    galaxies in each \tract{}.
